---
title: "Understanding Coordinate referencing system and reprojecting vector and raster data"
author: "himal2007"
date: "2021-06-02"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---


## Introduction

+ Loading libraries
```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(sf)
library(raster)
library(dplyr)
library(spData)
library(spDataLarge)
library(leaflet)
```

+ Function to find the epsg code and the UTM region for the geographic coordinates
```{r}
lonlat2UTM = function(lonlat) {
  utm = (floor((lonlat[1] + 180) / 6) %% 60) + 1
  if(lonlat[2] > 0) {
    utm + 32600
  } else{
    utm + 32700
  }
}

## example
epsg_utm_metema <-  lonlat2UTM(c(36.28284, 13.00668))
epsg_utm_metema
st_crs(epsg_utm_metema)$proj4string
```

+ Find more about the epsg code
```{r}
crs_codes = rgdal::make_EPSG()[1:4]
dplyr::filter(crs_codes, code == 32637)
st_crs(32637)$proj4string
```

### Transforming vector object to different CRS
+ Let's do this for the map of Ethiopia and also for the map of the world

#### Lambert azimuthal equal area

+ Visualize the world map
```{r}
plot(st_geometry(world))
st_crs(world)$proj4string
```

+ Tranform the world map into lambert azimuthal projection
```{r}
world_laea1 = st_transform(world, 
                           crs = "+proj=laea +x_0=0 +y_0=0 +lon_0=0 +lat_0=0")
plot(st_geometry(world_laea1))
st_crs(world_laea1)$proj4string
```

+ Tranform the world map into lambert azimuthal projection with Ethiopian centroid

+ Find the centroid for Ethiopia
```{r, message=FALSE}
m <- getData(name = "GADM", country = "ETH", level = 0)
m <- st_as_sf(m)
st_centroid(m) # 39.6262 8.629764
```

```{r}
world_laea2 = st_transform(world, 
                           crs = "+proj=laea +x_0=0 +y_0=0 +lon_0=0 +lat_0=0")
plot(st_geometry(world_laea2))
st_crs(world_laea2)$proj4string
```

### Reprojecting raster
+ Get altitude data for Ethiopia

```{r}
alt_eth <- getData(name = "alt", country = "ETH", mask = T)
crs(alt_eth)
plot(alt_eth)
```

#### Reprojecting to UTM CRS
```{r}
utm <- "+init=epsg:32637"
alt_eth_utm <- projectRaster(alt_eth, crs = utm, method = "bilinear")
crs(alt_eth_utm)
plot(alt_eth_utm)
```
+ method should be `ngb` for categorical raster

#### Reprojecting to Lambert Azimuthal Equal Area
```{r}
leaproj <- "+proj=laea +x_0=0 +y_0=0 +lon_0=39.62 +lat_0=8.62 +datum=WGS84 +units=km +no_defs" # you can set the unit to km

alt_eth_lea <- projectRaster(alt_eth, crs = leaproj, method = "bilinear")
crs(alt_eth_lea)
plot(alt_eth_lea)
```

+ Check the difference in the units of the coordinates


+ Reprojected rasters can be visualised with leaflet as well but the datum is discarded? I am not quite sure if the default projection is visualized or the tranformed projection.
```{r}
pal <- colorNumeric("RdYlBu", values(alt_eth_lea),
                    na.color = "transparent"
)

leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addRasterImage(alt_eth_lea, colors = pal, opacity = 0.5) %>%
  addLegend("bottomright",
            pal = pal, values = values(alt_eth_lea),
            title = "Altitude"
  ) %>%
  addScaleBar(position = c("bottomleft"))
```

